\section{Fundamentals}

\begin{defn}
    Two random vectors $(X_1, X_2, ..., X_n)$ and $(Y_1, Y_2, ..., Y_n)$ are said to be \textbf{independent} if $F(x_1, x_2, ..., x_m, y_1, y_2, ..., y_n) = F_1 (x_1, x_2, ..., x_m) F_2 (x_1, x_2, ..., x_n)$ for all $(x_1, x_2, ..., x_m, y_1, y_2, ..., y_n) \in \mathbb{R}^{m+n}$ where $F, F_1, F_2$ are the joint CDF's of $(X_1, X_2, ..., X_m, Y_1, Y_2, ..., Y_n)$, $(X_1, X_2, ..., X_m)$ and $(Y_1, Y_2, ..., Y_n)$ respectively.
\end{defn}

\begin{thm}
    Let $X = (X_1, X_2, · · · , X_m)$ and $Y = (Y_1, Y_2, · · · , Y_n)$ be independent random vectors. Then the component $X_j$ of $X(j = 1, 2, · · · , m)$ and the component $Y_k$ of $Y(k = 1, 2, · · · , n)$ are independent random variables. If $h$ and $g$ are Borel-measurable functions, $h(X_1, X_2, · · · , X_m)$ and $g(Y_1, Y_2, · · · , Y_n)$ are independent.
\end{thm}

\begin{thm}
    Let $X$ be a random variable with pdf $f(x)$. Then the pdf of $aX+b$ where $a \neq 0, b \in \mathbb{R}$ is given by $\displaystyle \frac{1}{a} f \left (\frac{x-b}{a} \right)$
\end{thm}

\begin{thm}
    If $X$ and $Y$ are independent continuous random variables with pdfs $f_X(x)$ and $f_Y(y)$, then the pdf of $Z = X+Y$ is $f_Z\left (z\right ) = \displaystyle\int_{- \infty}^\infty f_X \left (w\right ) f_Y \left (z - w\right ) \mathop{\mathrm{d} w}$
\end{thm}


\subsection{Expectation, Variance and Moments}

\begin{defn}
    The \textbf{expectation} of a random variable $X$ having values $x_1, x_2, ..., x_n$ is defined as $E\left (X\right ) = x_1  P\left (X = x_1\right ) + x_2 P\left (x = x_2\right ) + ... + x_n P\left (X = x_n\right ) =  \sum_{j = 1}^n x_j P\left (X = x_j\right ) = x_1 f\left (x_1\right ) + x_2 f\left (x_2\right ) + ... + x_n f\left (x_n\right ) = \sum_{j=1}^n x_j f\left (x_j\right )$ where $f$ is the distribution of $X$.
\end{defn}


\begin{samepage}
\begin{thm}
    The expectation has the following properties:

    \begin{enumerate}
        \item $E(c X) = c E(X)$ where $c$ is any constant
        \item If $X$ and $Y$ are any random variables then $E(X + Y) = E(X) + E(Y)$
        \item If $X$ and $Y$ are independent random variables, then $E(XY) = E(X)E(Y)$
    \end{enumerate}
\end{thm}
\end{samepage}

\begin{defn}
    The \textbf{variance} is defined as $\text{Var} (X) = \sigma ^2 = E[ \  (X - \mu)^2 \  ]$.
    The \textbf{standard deviation} is defined as $\sigma = \sqrt{\text{Var} (X)}$.
\end{defn}

\begin{samepage}
\begin{thm}
    The variance has the following properties:

    \begin{enumerate}
        \item $\sigma ^2 = E\left [ \left (X - \mu \right ) ^2 \right ] = E\left (X ^2\right ) - \mu ^2 = E\left (X ^2\right ) - \left [E\left (X\right )\right ] ^2$ where $\mu = E\left (X\right )$
        \item If $c$ is any constant, $\text{Var}(cX) = c^2 \text{Var}(X)$
        \item The quantity $E[ (X - a)^2 ]$ is a minimum where $a = \mu = E(X)$
        \item If $X$ and $Y$ are independent random variables, then \\ $\text{Var}(X+Y) = \text{Var}(X) + \text{Var}(Y)$ and $\text{Var}(X-Y) = \text{Var}(X) + \text{Var}(Y)$
    \end{enumerate}
\end{thm}
\end{samepage}

\begin{defn}
    The $r$-th \textbf{moment} of a random variable $X$ about the mean $\mu$ is defined as $\mu_r = E[\  (X - \mu)^r \ ]$ where $r=0, 1, 2, ...$.

    The $r$-th moment of $X$ about the origin is defined as $\mu_r ' = E(X^r)$.
\end{defn}

It follows that $\mu_0 = 0$, $\mu_1 = 1$, $\mu_2 = \sigma ^2$

\begin{defn}
    Let $X$ be a random variable defined on $\left (\Omega,  \mathcal{F}, P\right )$. The function $M\left (t\right ) = E\left [e ^{t X}\right ]$ is called the \textbf{moment generating function} (MGF) of the random variable $X$ if the expectation on the right side exists in some neighbourhood of the origin. If the expectation on the right side does not exist in any neighbourhood of the origin, then we say the MGF does not exist.
\end{defn}

\begin{samepage}
The $r$-th derivative of the moment generating funtion is the $r$-th moment about the origin $\mu_r '$.
\begin{thm}
    If the MGF $M(s)$ of a random variable $X$ exists, then the MGF $M(s)$ has derivatives of all orders at $s=0$ and \\

    $M^{(k)} (s) |_{s=0} = EX^k$ for positive integer $k$
\end{thm}

\end{samepage}
\begin{thm}
    The moment generating function has the following properties:
    \begin{enumerate}
        \item For any constants $a$ and $b$, the mgf of the random variable $a X + b$ is given by $M_{aX+b} = e^{bt} M_X(at)$
        \item If $X$ and $Y$ are independent random variables having moment generating functions $M_X(t)$ and $M_Y(t)$ respectively, then $M_{X+Y}(t) = M_X(t) M_Y(t)$
        \item \textbf{Uniqueness Theorem} Suppose that $X$ and $Y$ are random variables having moment generating functions $M_X(t)$ and $M_Y(t)$ respectively. Then $X$ and $Y$ have the same probability distribution if and only if $M_X(t) = M_Y(t)$ identically.
    \end{enumerate}
\end{thm}

\begin{defn}
    Let $X_1, X_2, ..., X_n$ be a jointly distributed or $(X_1, X_2, ..., X_n)$ be a random vector. If $E[\exp ( \sum_{j=1}^n t_j X_j )]$ exists for $|t_j| \leq h_j$, $j = 1, 2, ...,n$, we write $M(t_1, t_2, ..., t_n) = E[\exp(t_1 X_1 + t_2 X_2 + ... + t_n X_n)]$ and call it the MGF of $X_1, X_2, ..., X_n$ or simply, the \textbf{joint moment generating function} (joint MGF) of the random vector $(X_1, X_2, ..., X_n)$
\end{defn}

\begin{thm}
    The joint MGF $M(t_1, t_2)$ uniquely determines the joint distribution of $(X, Y)$.
    Conversely, if the joint MGF exists it is unique.
\end{thm}

\begin{thm}
    The joint MGF $M(t_1, t_2)$ completely determines the marginal distributions of $X$ and $Y$.

    $M(t_1, 0) = E[\exp(t_1 X)] = M_X (t_1)$ and $M(0, t_2) = E[\exp(t_2 X)] = M_Y(t_2)$
\end{thm}

\begin{thm}
    $X$ and $Y$ are independent random variables if and only if $M(t_1, t_2) = M(t_1, 0) M(0, t_2)$ for all $t_1 \in [-h_1, h_1], t_2 \in [-h_2, h_2]$
\end{thm}

\subsection{Distributions}

\begin{defn}
    Let $p$ be the probability that an event will happen in any single Bernoulli trial (trial with outcomes either success or failure). The probability that an event will happen exactly $x$ times in $n$ trials is given by the \textbf{Binomial Random Variable} with pmf distribution $f(x) = {n \choose k} p^x q^{n-x} = \displaystyle \frac{n!}{x! (n-x)!} p^x q^{n-x}$, where $x = 0, 1, ... , n$.
\end{defn}

The binomial random variable has mean $\mu = np$ and variance $\sigma^2 = npq$.

\begin{defn}
    The \textbf{Poisson Random Variable} has pmf distribution $f(x) = \displaystyle \frac{\lambda^x e^{-\lambda}}{x!}$, $x = 0, 1, 2, ...$ and $\lambda > 0$.
\end{defn}

The Poisson Random Variable has mean $\mu = \lambda$ and variance $\sigma^2 = \lambda$.
\begin{prop}
    The Poisson Random Variable with $\lambda = np$ is the limiting case of the Binomial Distribution. It approximates the Binomial Random variable $\text{Binomial}(n,p)$ when $n$ is large and probability of occurrence of an event $p$ is close to $0$.
\end{prop}

\begin{defn}    
    The \textbf{uniform Random Variable} has the pdf distribution $f(x) = \begin{cases} \frac{1}{b-a} & a \leq x \leq b \\ 0 & \text { otherwise }\end{cases}$.
\end{defn}

The uniform random variable has mean $\mu = \frac{a + b}{2}$ and variance $\sigma^2 = \frac{1}{12} (b-a)^2$.

\begin{defn}
    The \textbf{Normal Random Variable}, also known as the gaussian random variable, has pdf distribution $f(x) = \frac{1}{\sigma \sqrt{2 \pi}}  \exp\left(\displaystyle \frac{-(x-\mu)^2}{2 \sigma^2}\right)$ where $-\infty < x < \infty$ where $\mu$ and $\sigma$ are the mean and standard deviation respectively.
\end{defn}

When $\mu = 0$ and $\sigma = 1$, we get the \textbf{standard normal random variable} with distribution $f(x) = \displaystyle\frac{1}{2\pi} \exp(-\frac{x^2}{2})$.

\begin{prop}
    When $n$ is large and neither $p$ or $q$ is too close to $0$, the binomial random variable $X$ can be approximated by a normal distribution with mean $np$ and standard deviation $\sqrt{npq}$. The approximation is very good when $np, nq > 5$.
\end{prop}

\begin{prop}
    The Poisson Distribution approaches the normal distribution $N(\lambda, \sqrt{\lambda})$ as $\lambda \to \infty$, i.e. the Poisson distribution is asymptotically normal.
\end{prop}

\begin{defn}
    A variable the pdf $f(x) = \displaystyle \frac{1}{\sigma \pi (1 + \left( \frac{x - \mu}{\sigma} \right)^2 )}$, $x \in \mathbb{R}$ where $\sigma > 0, \mu \in \mathbb{R}$ is called a \textbf{Cauchy random variable} with parameter $\mu$ and $\sigma^2$.
    We write $X \sim \mathcal{C}(\mu, \sigma^2)$ for Cauchy random $X$ with pdf.
\end{defn}

The Cauchy random variable centered at $0$ (i.e. $\mu = 0$ is given by the distribution $f(x) = \displaystyle \frac{\sigma}{\pi ( \sigma^2 + x^2}$ where $\sigma > 0, -\infty < x < \infty$.

The mean, variance, higher moments, moment generating function of a Cauchy random variable do not exist.

\begin{defn}
    $\Gamma(\alpha)$ denotes the gamma function which is defined as $\Gamma(\alpha) = \displaystyle\int_0^\infty t ^{\alpha - 1} e ^{-t} \mathop{\mathrm{d} t}$, where $\alpha > 0$. It is a commonly used extension of the factorial function.
\end{defn}

\begin{thm}
    The Gamma function has the following properties:
    \begin{enumerate}
        \item The improper integral converges for all $\alpha > 0$.
        \item $\Gamma(\alpha) > 0$  for all $\alpha > 0$, $\Gamma(\alpha) \to \infty$ as $\alpha \to 0$.
        \item $\Gamma(1) = 1$
        \item $\Gamma(\alpha + 1) = \alpha \Gamma(\alpha)$
        \item For $n \in \mathbb{N}$, $\Gamma(n) = (n-1)!$
        \item $\Gamma(\frac{1}{2}) = \sqrt{\pi}$
    \end{enumerate}
\end{thm}

\begin{defn}
    A random variable with the pdf $f(x) = \displaystyle  \frac{1}{\Gamma(\alpha) \beta^\alpha} x^{\alpha - 1} e^{-x/\beta}$ where $0 < x < \infty, \alpha > 0, \beta > 0$ is called a \textbf{gamma random variable} $G(\alpha, \beta)$ with parameters $\alpha$ and $\beta$.
\end{defn}
$G(1, \beta) = \text{exp}(\beta)$ with pdf $f(x) = \frac{1}{\beta} e^{-x/\beta}$, $x > 0$.

The gamma distribution has mean $\mu = \alpha \beta$ and variance $\sigma^2 = \alpha \beta^2$.

\begin{defn}
    The \textbf{chi-square distribution} is \\
    $f(x) = \displaystyle \frac{1}{\Gamma(\frac{p}{2}) 2^{\frac{p}{2}}} x^{\frac{p}{2} - 1} e ^{-x/2}$ where $0 < x < \infty$.

    \vspace{0.5cm}
    $\chi_p^2$ denotes a chi-square random variable with $p$ degrees of freedom. It is a special case of the gamma distribution.
\end{defn}

The chi-square distribution is a special case of the gamma distribution with $\alpha = \frac{p}{2}$ and $\beta = 2$.

The mean of the chi-square distribution is given by $\mu = p$ and the variance is given by $\sigma^2 = 2p$.

\begin{prop}
Let $X \sim N(0,1)$. Then $X^2 \sim \chi_1^2$
\end{prop}

\begin{prop}
    Let $X_1, X_2, ..., X_n$ be independent normal random variables with mean $0$ and variance $1$. Then $\chi^2 = X_1^2 + X_2^2 + ... + X_p^2$ is chi-square distributed with $p$ degrees of freedom.
\end{prop}
